# Ethiopia

Relevant sources of household-level micro records for Ethiopia include:

* 2011/12 ERSS (**ESS Wave 1**)  
* 2013/14 Ethiopia Socioeconomic Surveys (**ESS Wave 2**)  
* 2015/16 Ethiopia Socioeconomic Surveys (**ESS Wave 3**)  
* 2012/13 Ethiopia’s Agricultural Sample Survey (AgSS)  
* 2012 ATA Baseline Survey (3,000 hhlds), see [@minot2013agataeth].

The ERSS 2011/12 sample is designed to be representative of **rural** and **small town areas** of Ethiopia. It is a sub-sample of the 2011/2012 Agricultural Sample Survey (AgSS). The small town sample comes from the universe of small town EAs. The ERSS sample size provides estimates at the national level for rural and small town households. At the regional level, it provides estimates for **four regions** including Amhara, Oromiya, SNNP, and Tigray. The sample is a two-stage probability sample. The first stage of sampling entailed selecting primary sampling units, which are a sample of the CSA enumeration areas (EAs). For the rural sample, 290 EAs were selected from the AgSS EAs. The AgSS EAs were selected based on probability proportional to size of the total EAs in each region. For small town EAs, a total of 43 EAs were selected. In order to ensure sufficient sample in the most populous regions (Amhara, Oromiya, SNNP, and Tigray), quotas were set for the number of EAs in each region. The sample is not representative for each of the small regions including Afar, Benshangul Gumuz, Dire Dawa, Gambella, Harari, and Somali regions. However, estimates can be produced for a combination of all smaller regions as one "other regions" category.

The 2nd wave ESS 2013/14 adds samples from large town areas. The population frame for the urban expansion consists of all households in towns with population greater than 10,000 people. This population cut off of 10,000 people is the same threshold that is used to define small towns. This wave is nationally representative. The existing panel data (2011/12-2013/14) is only for rural and small towns. The ESS sample size provides estimates at the national level for rural and small town households. At the regional level, it provides estimates for five regions including Addis Ababa, Amhara, Oromiya, SNNP, Tigray, and "other regions".

Note that FAO RIGA-H database is not yet available for ESS Wave 3 (as of Nov 8, 2017).

## Household Sampling and Stratification

```{r eth-samples}

tmp <- eth[, .(
  Regions = uniqueN(region),
  Districts = uniqueN(paste(region, district_cd)),
  Sample = .N,
  Rural = sum(rural=="rural", na.rm=T),
  Urban = sum(rural=="urban", na.rm=T),
  `Farm hhlds` = sum(farm==TRUE)
  ), keyby=.(Survey=svyCode)]

  # Population = sum(weight, na.rm=T),
  # Rural = sum(c(rural=="rural") * weight, na.rm=T),
  # Urban = sum(c(rural=="urban") * weight, na.rm=T)

tmp[, `:=`(
  `Rural (pct)` = 100*Rural/Sample,
  `Urban (pct)` = 100*Urban/Sample,
  `Farm hhlds (pct)` = 100*`Farm hhlds`/Sample
)]

kable(tmp,
  justify="ccrrrrrrrrrrrr", format.args=list(big.mark=","), digits=0,
  caption="Composition of Ethiopia ESS Harmonized Survey Samples")

```

```{r eth-strata, fig.cap="(Appoximate) Household Locations and Stratification across Survey Waves, Ethiopia", fig.height=3}

# Map household by strata to verify
tm_shape(g2[g2$svyCode=="eth-erss-2011",]) + tm_polygons("grey90", border.col="grey50") +
  tm_shape(eth.pts) + 
  tm_dots("strata", size=0.08, palette=viridis(11), title="Survey\nStratification") + 
  tm_facets("survey", ncol=2) + tm_layout(legend.outside=T)

```

```{r, eth-panels}

# Panels
tmp <- dcast(eth, survey~pcount)
setnames(tmp, c("Survey", "1 visit", "2 visits"))
kable(tmp,
  justify="lrr", format.args=list(big.mark=","), digits=0,
  caption="Attrition of Households across Panels, Ethiopia")

```

Out of the *`r prty(eth[svyCode=="eth-erss-2011", .N])` households* recorded in the 2011 ERSS *`r prty(tmp[1,3])` households* were re-visited in all 2 waves, while *`r prty(tmp[1,2])`* out of `r prty(eth[svyCode=="eth-ess-2013", .N])` households in the second ESS 2013 sample were only interviewed once. This is due to attrition and tracking of split households across waves.

## Distribution of Farm Holdings across Farm Sizes and Farm Types

The plots below show the distribution of key variables in the unweighted sample of households across survey waves. The median line is in red, mean line in green, and the blue region is the inferred 95% confidence interval of the mean.

### Land Holding

```{r eth-land, fig.cap="Trends in Farm Sizes across Survey Waves (ha)"}

par(mfrow=c(1,2))
pplot(croparea_imp~survey, eth[farm==TRUE & croparea_4ha=="<= 4 ha"],
  xlab="Farms below 4 ha", ylab="Farm Size (ha)",
  note=eth[, 100*sum(is.na(croparea_imp)|croparea_imp==0)/.N])

pplot(croparea_imp~survey, eth[farm==TRUE & croparea_4ha=="> 4 ha"],
  xlab="Farms above 4 ha", ylab="Farm Size (ha)", ylim=c(4,25),
  note=eth[, 100*sum(is.na(croparea_imp)|croparea_imp==0)/.N])

```



### Trends in Commercialization

```{r eth-cc, fig.cap="Crop Commercialization across Survey Waves (percent of crop production)"}

par(mfrow=c(1,2))
pplot(100*cropsales_sh~croparea_4ha+survey, eth[farm==TRUE],
  ylab="Crop Comm. (vol., source: authors)",
  note=eth[, 100*sum(is.na(cropsales_sh)|cropsales_sh==0)/.N])

pplot(100*shcropsold~croparea_4ha+survey, eth,
  ylab="Crop Comm. (val., source: RIGA)",
  note=eth[, 100*sum(is.na(cropsales_sh)|cropsales_sh==0)/.N])

```

```{r eth-ac, fig.cap="Agricultural Commercialization across Survey Waves (percent of ag. production)"}

pplot(100*agsales_sh~croparea_4ha+survey, eth[farm==TRUE],
  ylab="Agricultural Commercialization",
  note=eth[, 100*sum(is.na(agsales_sh)|agsales_sh==0)/.N])

```

### Trends in Income Diversification

The shares of farm and non-farm income are derived from RIGA components, namely:

```
aggross = totagprod + farmrntimp + agr_wge+ ifelse(selfimp1 < 0, NA, selfimp1)
naggross = transfersgrossimp + otherincimp + nonfarmrntimp + nonagr_wge + ifelse(selfemp < 0, NA, selfemp)
```

```{r eth-div, fig.cap="Non-Farm Income (percent of gross income)"}

pplot(100*naggross_sh~croparea_4ha+survey, eth[farm==TRUE],
  ylab="Non-Farm Income Share",
  note=eth[, 100*sum(is.na(naggross_sh)|naggross_sh==0)/.N])
```


### 3x3 Farm Classification

```{r eth-prop9, fig.cap="Est. Proportions of Farm Holdings across Farm Sizes and Categories", fig.height=3}

par(mfrow=c(1,2), mar=c(3,3,0,.25))
# plot(svytable(~croparea_3clas+class9, eth.svy[["eth3"]]), main=NA, las=1, col=viridis(9),
#   ylab=NA, xlab="Typology 9-class (2012/13)")
plot(svytable(~croparea_3clas+class9, eth.svy.farm[["eth2"]]), main=NA, las=1, col=viridis(9),
  ylab=NA, xlab="Typology 9-class (2013/14)")
plot(svytable(~croparea_3clas+class9, eth.svy.farm[["eth1"]]), main=NA, las=1, col=viridis(9),
  ylab=NA, xlab="Typology 9-class (2011/12)")

```

```{r eth-proptab93, results="asis", eval=FALSE}

# ESS3
tmp <- svyCrossProp(c("croparea_3clas", "cropsales_3clas", "naggross_3clas"), eth.svy.farm[["eth3"]])

table_options(HTMLcaption="(#tab:proptab93) Est. Proportions of Farm Holdings across Categories (ESS, 2015/16, percent)")
html(tabular(Heading("Income Diversification")*naggross_3clas*(mean+CI)~Heading("Crop Commercialization")*croparea_3clas*Heading()*cropsales_3clas*Heading()*identity,
  data=tmp), rmarkdown=T)

```

```{r eth-proptab92, results="asis"}

# ESS2
tmp <- svyCrossProp(c("croparea_3clas", "cropsales_3clas", "naggross_3clas"), eth.svy.farm[["eth2"]])

table_options(HTMLcaption="(#tab:proptab92) Est. Proportions of Farm Holdings across Categories (ESS, 2013/14, percent)")
html(tabular(Heading("Income Diversification")*naggross_3clas*(mean+CI)~Heading("Crop Commercialization")*croparea_3clas*Heading()*cropsales_3clas*Heading()*identity,
  data=tmp), rmarkdown=T)

```

```{r eth-proptab91, results="asis"}
# ERSS1
tmp <- svyCrossProp(c("croparea_3clas", "cropsales_3clas", "naggross_3clas"), eth.svy.farm[["eth1"]])

table_options(HTMLcaption="(#tab:proptab91) Est. Proportions of Farm Holdings across Categories (ERSS, 2011/12, percent)")
html(tabular(Heading("Income Diversification")*naggross_3clas*(mean+CI)~Heading("Crop Commercialization")*croparea_3clas*Heading()*cropsales_3clas*Heading()*identity,
  data=tmp), rmarkdown=T)

```

Proportions of **small holdings** across categories (farms below 4 ha).

```{r eth-proptab59b, results="asis"}

# To ease comparisons with EPAR
tmp1 <- lapply(eth.svy.shf, function(x) svyCrossProp(c("naggross_3clas", "cropsales_3clas"), x))
tmp <- rbindlist(tmp1)
tmp2 <- c("2013/2014", "2011/2012")
tmp[, year := rep(tmp2, each=9)]
tmp[, year := factor(year, levels=tmp2)]

table_options(HTMLcaption="(#tab:proptab59b) Est. Proportions of Farm Holdings across Categories (ESS all waves, percent, farms below 4 ha)")
html(tabular((`Year`=year)*Heading("Income Diversification")*naggross_3clas*(mean+CI)~Heading("Crop Commercialization")*cropsales_3clas*Heading()*identity,
  data=tmp), rmarkdown=T)

```


### Simplified Farm Classification (5-class)

```{r eth-prop5, fig.cap="Est. Proportions of Farm Holdings across Farm Sizes and Categories"}

par(mfrow=c(1,2), mar=c(3,3,0,.25))
# plot(svytable(~croparea_3clas+class5, eth.svy[["eth3"]]), main=NA, las=1, col=viridis(5),
#   ylab=NA, xlab="Typology 5-class (2015/16)")
plot(svytable(~croparea_3clas+class5, eth.svy[["eth2"]]), main=NA, las=1, col=viridis(5),
  ylab=NA, xlab="Typology 5-class (2013/14)")
plot(svytable(~croparea_3clas+class5, eth.svy[["eth1"]]), main=NA, las=1, col=viridis(5),
  ylab=NA, xlab="Typology 5-class (2011/12)")

```

```{r eth-proptab56, results="asis", eval=FALSE}

# ESS3
tmp <- svyCrossProp(c("croparea_3clas", "class5"), eth.svy.farm[["eth3"]])

table_options(HTMLcaption="(#tab:proptab56) Est. Proportions of Farm Holdings across Categories (ESS, 2015/16, percent)")
html(tabular(Heading("Farm Size")*croparea_3clas*(mean+CI)~Heading("Farm Types")*class5*Heading()*identity,
  data=tmp), rmarkdown=T)

```

```{r eth-proptab55, results="asis"}

# ESS2
tmp <- svyCrossProp(c("croparea_3clas","class5"), eth.svy.farm[["eth2"]])

table_options(HTMLcaption="(#tab:proptab55) Est. Proportions of Farm Holdings across Categories (ESS, 2013/14, percent)")
html(tabular(Heading("Farm Size")*croparea_3clas*(mean+CI)~Heading("Farm Types")*class5*Heading()*identity,
  data=tmp), rmarkdown=T)

```

```{r eth-proptab54, results="asis"}

# ERSS1
tmp <- svyCrossProp(c("croparea_3clas", "class5"), eth.svy.farm[["eth1"]])

table_options(HTMLcaption="(#tab:proptab54) Est. Proportions of Farm Holdings across Categories (ERSS, 2011/12, percent)")
html(tabular(Heading("Farm Size")*croparea_3clas*(mean+CI)~Heading("Farm Types")*class5*Heading()*identity,
  data=tmp), rmarkdown=T)

```

Proportions of **small holdings** across categories (farms below 4 ha)


```{r eth-proptab56b, results="asis"}

# To ease comparisons with EPAR
tmp1 <- lapply(eth.svy.shf, function(x) svyCrossProp("class5", x))
tmp <- rbindlist(tmp1)
tmp2 <- c("2013/2014", "2011/2012")
tmp[, year := rep(tmp2, each=5)]
tmp[, year := factor(year, levels=tmp2)]

table_options(HTMLcaption="(#tab:proptab56) Est. Proportions of Farm Holdings across Categories (TNPS all waves, percent, farms below 4 ha)")
html(tabular((`Year`=year)*(mean+CI)~Heading("Farm Types")*class5*Heading()*identity,
  data=tmp), rmarkdown=T)

```


## Farm Segmentation along Farm Types (5-class) and Agricultural Potential Quadrants

```{r eth-figquad, fig.cap="Agricultural Potential Quadrants, Ethiopia"}

tmp <- g2[g2$svyCode=="eth-ess-2013",]
tm_shape(mask(rquad, tmp)) + tm_raster(pal=viridis(4), n=4, labels=quad,
  title="Agricultural\nPotential\nQuadrants, \nEthiopia") +
  tm_shape(tmp, is.master=T) + tm_borders("grey", lwd=.4) +
  tm_layout(legend.outside=T)

```

### All Farms

```{r eth-propquad5, fig.cap="Est. Proportions of Farm Holdings across Farm Sizes and Quadrants (all farm sizes combined)", dev.args=list(pointsize=14)}

par(mfrow=c(1,2), mar=c(3,0.25,0.25,.25))
# plot(svytable(~class5+seg_quad, eth.svy[["eth3"]]), main=NA, las=2, col=viridis(4),
#   ylab=NA, xlab="5-types and Quadrants (2012/13)")
plot(svytable(~class5+seg_quad, eth.svy[["eth2"]]), main=NA, las=2, col=viridis(4),
  ylab=NA, xlab="5-types and Quadrants (2013/14)")
plot(svytable(~class5+seg_quad, eth.svy[["eth1"]]), main=NA, las=2, col=viridis(4),
  ylab=NA, xlab="5-types and Quadrants (2011/12)")

```

```{r eth-propquadtab53, results="asis", eval=FALSE}

# ESS3
tmp <- svyCrossProp(c("seg_quad", "class5"), eth.svy.farm[["eth3"]])

table_options(HTMLcaption="(#tab:propquadtab53) Est. Proportions of Farm Holdings across Types and Quadrants (TNPS3, 2012/13, percent)")
html(tabular(Heading("Quadrant")*seg_quad*(mean+CI)~Heading("Farm Types")*class5*Heading()*identity,
  data=tmp), rmarkdown=T)

```

```{r eth-propquadtab52, results="asis"}

# ESS2
tmp <- svyCrossProp(c("seg_quad", "class5"), eth.svy.farm[["eth2"]])

table_options(HTMLcaption="(#tab:propquadtab52) Est. Proportions of Farm Holdings across Types and Quadrants (ESS2, 2013/2014, percent)")
html(tabular(Heading("Quadrant")*seg_quad*(mean+CI)~Heading("Farm Types")*class5*Heading()*identity,
  data=tmp), rmarkdown=T)

```

```{r eth-propquadtab51, results="asis"}

# ERSS1
tmp <- svyCrossProp(c("seg_quad", "class5"), eth.svy.farm[["eth1"]])

table_options(HTMLcaption="(#tab:propquadtab51) Est. Proportions of Farm Holdings across Types and Quadrants (ERSS, 2011/2012, percent)")
html(tabular(Heading("Quadrant")*seg_quad*(mean+CI)~Heading("Farm Types")*class5*Heading()*identity,
  data=tmp), rmarkdown=T)

```

### Small Farms (below 4 ha)

```{r eth-propquad5sh, fig.cap="Est. Proportions of Farm Holdings across Farm Sizes and Quadrants (farms below 4 ha)", dev.args=list(pointsize=14)}

par(mfrow=c(1,2), mar=c(3,0.25,2,.25))
# plot(svytable(~class5+seg_quad, eth.svy.shf[["eth3"]]), main=NA, las=2, col=viridis(4),
#   ylab=NA, xlab="5-types and Quadrants (2012/13)")
plot(svytable(~class5+seg_quad, eth.svy.shf[["eth2"]]), main=NA, las=2, col=viridis(4),
  ylab=NA, xlab="5-types and Quadrants (2013/14)")
plot(svytable(~class5+seg_quad, eth.svy.shf[["eth1"]]), main=NA, las=2, col=viridis(4),
  ylab=NA, xlab="5-types and Quadrants (2011/12)")

```

```{r eth-propquadtab53sh, results="asis", eval=FALSE}

# ESS3
tmp <- svyCrossProp(c("seg_quad", "class5"), eth.svy.shf[["eth3"]])

table_options(HTMLcaption="(#tab:propquadtab53sh) Est. Proportions of Farm Holdings across Types and Quadrants (ESS, 2015/16, percent, farms below 4 ha)")
html(tabular(Heading("Quadrant")*seg_quad*(mean+CI)~Heading("Farm Types")*class5*Heading()*identity,
  data=tmp), rmarkdown=T)

```

```{r eth-propquadtab52sh, results="asis"}

# ESS2
tmp <- svyCrossProp(c("seg_quad", "class5"), eth.svy.shf[["eth2"]])

table_options(HTMLcaption="(#tab:propquadtab52sh) Est. Proportions of Farm Holdings across Types and Quadrants (ESS, 2013/2014, percent, farms below 4 ha)")
html(tabular(Heading("Quadrant")*seg_quad*(mean+CI)~Heading("Farm Types")*class5*Heading()*identity,
  data=tmp), rmarkdown=T)

```

```{r eth-propquadtab51sh, results="asis"}

# ERSS
tmp <- svyCrossProp(c("seg_quad", "class5"), eth.svy.shf[["eth1"]])

table_options(HTMLcaption="(#tab:propquadtab51sh) Est. Proportions of Farm Holdings across Types and Quadrants (ERSS, 2011/2012, percent, farms below 4 ha)")
html(tabular(Heading("Quadrant")*seg_quad*(mean+CI)~Heading("Farm Types")*class5*Heading()*identity,
  data=tmp), rmarkdown=T)

```



