# Ethiopia

Relevant sources of household-level micro records for Ethiopia include:

* 2011/12 ERSS (**ESS Wave 1**)  
* 2013/14 Ethiopia Socioeconomic Surveys (**ESS Wave 2**)  
* 2015/16 Ethiopia Socioeconomic Surveys (**ESS Wave 3**)  
* 2012/13 Ethiopia’s Agricultural Sample Survey (AgSS)  
* 2012 ATA Baseline Survey (3,000 hhlds), see [@minot2013agataeth].

The ERSS 2011/12 sample is designed to be representative of **rural** and **small town areas** of Ethiopia. It is a sub-sample of the 2011/2012 Agricultural Sample Survey (AgSS). The small town sample comes from the universe of small town EAs. The ERSS sample size provides estimates at the national level for rural and small town households. At the regional level, it provides estimates for **four regions** including Amhara, Oromiya, SNNP, and Tigray. The sample is a two-stage probability sample. The first stage of sampling entailed selecting primary sampling units, which are a sample of the CSA enumeration areas (EAs). For the rural sample, 290 EAs were selected from the AgSS EAs. The AgSS EAs were selected based on probability proportional to size of the total EAs in each region. For small town EAs, a total of 43 EAs were selected. In order to ensure sufficient sample in the most populous regions (Amhara, Oromiya, SNNP, and Tigray), quotas were set for the number of EAs in each region. The sample is not representative for each of the small regions including Afar, Benshangul Gumuz, Dire Dawa, Gambella, Harari, and Somali regions. However, estimates can be produced for a combination of all smaller regions as one "other regions" category.

The 2nd wave ESS 2013/14 adds samples from large town areas. The population frame for the urban expansion consists of all households in towns with population greater than 10,000 people. This population cut off of 10,000 people is the same threshold that is used to define small towns. This wave is nationally representative. The existing panel data (2011/12-2013/14) is only for rural and small towns. The ESS sample size provides estimates at the national level for rural and small town households. At the regional level, it provides estimates for five regions including Addis Ababa, Amhara, Oromiya, SNNP, Tigray, and "other regions".

Note that FAO RIGA-H database is not yet available for ESS Wave 3 (as of Nov 8, 2017).

```{r eth-samples}

tmp <- eth[, .(
  Regions = uniqueN(svyL1Nm),
  Districts = uniqueN(paste(svyL1Cd, svyL2Cd)),
  Sample = .N,
  Population = sum(weight, na.rm=T),
  Rural = sum(c(rural=="rural") * weight, na.rm=T),
  Urban = sum(c(rural=="urban") * weight, na.rm=T)
), keyby=.(Survey=svyCode)]

tmp[, `:=`(
  `Rural (pct)` = 100*Rural/Population,
  `Urban (pct)` = 100*Urban/Population
)]

kable(tmp,
  justify="ccrrrrrrrrrrr", format.args=list(big.mark=","), digits=0,
  caption="Composition of Ethiopia ESS Harmonized Survey Samples")

```

```{r eth-strata, fig.cap="(Appoximate) Household Locations and Stratification across Survey Waves, Ethiopia", fig.height=3}

# Map household by strata to verify
tm_shape(g2[g2$svyCode=="eth-erss-2011",]) + tm_polygons("grey90", border.col="grey50") +
  tm_shape(eth.pts) + 
  tm_dots("strata", size=0.08, palette=viridis(11), title="Survey\nStratification") + 
  tm_facets("survey", ncol=2) + tm_layout(legend.outside.size=.1)

```

```{r, eth-panels}

# Panels
tmp <- dcast(eth, survey~pcount)
setnames(tmp, c("Survey", "1 visit", "2 visits"))
kable(tmp,
  justify="lrr", format.args=list(big.mark=","), digits=0,
  caption="Attrition of Households across Panels, Ethiopia")

```

Out of the *`r prty(eth[svyCode=="eth-erss-2011", .N])` households* recorded in the 2011 ERSS *`r prty(tmp[1,3])` households* were re-visited in all 2 waves, while *`r prty(tmp[1,2])`* out of `r prty(eth[svyCode=="eth-ess-2013", .N])` households in the second ESS 2013 sample were only interviewed once. This is due to attrition and tracking of split households across waves.

## Farm Segmentation along Farm Types (5-class) and Agricultural Potential Quadrants

```{r eth-figquad, fig.cap="Agricultural Potential Quadrants, Ethiopia"}

tmp <- g2[g2$svyCode=="eth-ess-2013",]
tm_shape(mask(rquad, tmp)) + tm_raster(pal=viridis(4), n=4, labels=quad,
  title="Agricultural\nPotential\nQuadrants, \nEthiopia") +
  tm_shape(tmp, is.master=T) + tm_borders("grey", lwd=.4) +
  tm_layout(legend.outside=T)

```

## Distribution of Farm Holdings across Farm Sizes and Farm Types

The plots below show the distribution of key variables in the unweighted sample of households across survey waves. The median line is in red, mean line in green, and the blue region is the inferred 95% confidence interval of the mean.

[pending]


